---
title: "Recommended measures to reduce business flights at Eawag"
author: "Flyaware interest group"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
    # css: my_css.css
  tint::tintPdf:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
subtitle: Final report to the directorate
bibliography: skeleton.bib
---
  
```{r setup, include=FALSE}
library(tint)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```

```{r setup data, include=FALSE}
source("read_in_clean_data.R")
source("utility_functions.R")
```


# A call for action

> "It is critical to immediately begin reducing net CO2 emissions and to eliminate them to zero worldwide between 2040 and 2050 at the latest."
>
> `r tufte::quote_footer('--- Scientists for future')`

This statement from the ‘Scientists for Future’^[https://www.scientists4future.org/statement-en-und-es/]  initiative was signed by 44 out of 87 (50.6%) Eawag group leaders and department heads. The director and deputy director of Eawag have signed this statement as well. Thus, the majority of the Eawag leadership is in agreement with this statement and is aware of the urgency of taking measures to reduce emissions.

One of the most effective ways for scientific institutions to reduce their climate impact is a drastic reduction of business flights. More than half of greenhouse gas emissions (GHGs) of Eawag's sister institution ETH Zurich are caused by business trips. Of those **93%** are due to business flights ^[https://ethz.ch/flugreisen]. Technical alternatives to substantially reduce GHG emissions of airplanes are not foreseeable for the near future. It is thus of immediate concern to reduce the number of flights.

Eawag is a research institution recognized worldwide and therefore has the potential to act as a role model for other research institutes by taking on its institutional responsibility for further reducing CO2 emissions. The time to be a rolemodel is now - academia and the world in general is finally waking up to the climate impacts of flying ^[for a helpful FAQ and original 2015 petition of academics worldwide see https://academicflyingblog.wordpress.com/]

This report is a call to the Eawag leadership to act. We propose nine concrete sets of measures to reduce Eawag’s CO2 emissions related to business flights. To aid in decision-making, We provide a list of highly accepted and highly impactful measures as priority targets for implementation.

```{r fig-margin, fig.margin = TRUE, fig.cap = "Employees of Eawag overwhelmingly support Eawag taking measures to aid in reducing business flights", fig.width=3.5, fig.height=3.5, cache=TRUE, message=FALSE, echo=FALSE}
plot_var_dist("should_eawag_support_flyless",y_limit = max_n,plot_xlabs = TRUE, plot_legend = FALSE, fill_var = "none", graph_title = "Eawag should support flying less")
```

We conducted a survey among all Eawag employees, in which roughly half of all employees (`r nrow(responses)`) participated. Employees were asked to evaluate the set of measures we present in this report. Given the high response rate, our results are highly likely to adequately reflect the views of Eawag employees on the topic of flying and the proposed measures.

The scientific evidence for man-made climate change and its dangers is robust. We believe that you as (environmental) scientists and members of the directorate want to overcome the knowledge – action gap and support change here at Eawag.

## What measures should Eawag take?

We suggest the following list of measures for Eawag to take right away. ^[The complete set of measures as presented to Eawag employees in the survey can be found at the end of this report. Clicking on individual measures links to the specific measure.]

- *[Official Statement](#official) (Eawag Directive)*

- [*Intranet page*](#intranet)

- [*Showcasing video conferences*](#showcase_video)

- [*Train travel as official policy*](#train_travel_official_policy)

- [*Compensation of train travel*](#train_travel_compensate)

- [*Compensate all flights*](#compensate_all_flighs)

- [*Double the compensation*](#double_compensation)

- [*Internal reduction targets*](#internal_competition)

- [*Improved database*](#database)


One of the most effective ways to reduce GHG emissions from flying is the reduction of long-distance flights ^[see eg. Jäckle, S. Eur Polit Sci (2019) 18: 630. https://doi.org/10.1057/s41304-019-00220-6]. Thus, we strongly recommend to prioritize a measure acting on long-distance flights. However, since the quite stringent measure we initially proposed was very controversial among Eawag employees, we propose the directorate adapt the measure to improve acceptance.

- [*Limit long-distance flights*](#overseas_restriction)


## Categorization of measures

To select the measures we propose for prioritized implementation we categorized them based on their acceptance among Eawag employees and their expected impact in terms of reducing GHG emissions. Acceptance was split into three groups, based on the distribution of acceptance^[However, it needs to be kept in mind that no single measure was rejected by a majority of survey respondents] among Eawag employees^[The distribution of variables can be explored interactively at: https://marioangst.shinyapps.io/flyaware_survey/ 
Survey data and analytical procedures can be accessed at https://github.com/marioangst/flyaware_survey]. Expected impact and expected costs are based on an estimate within the flyaware group and as such only represent an initial prior. We welcome further discussion on these points.

We suggest to categorize measures into four groups:

- High acceptance: These measures are universally supported by Eawag employees. They are mostly relatively low-cost. We recommend their immediate adoption.
- Very disputed: The most controversial, yet generally most impactful measures. We recommend to forego these for the moment, except for the measure on limiting [long-distance flights](#overseas_restriction). Here we recommend to adapt the relatively stringent measure we propose.
- Disputed, medium impact: Somewhat disputed, medium-impact measures. We recommend the adoption of these measures.
- Disputed, low impact: Somewhat disputed measures with relatively little impact. We do recommend not to prioritize these measures, except concerning the adoption of a measure [improving the database](#database) as it is the foundation for some other measures.

```{r, fig.width = 9, fig.height = 5, fig.cap = "Categorization of measures proposed based on acceptance, expected impact and expected cost. Each circle represents a measure, sized by expected costs.", warning=FALSE, message=FALSE, echo=FALSE}
library(ggrepel)

pos <- position_jitter(0.2,0.2,seed = 1)

ggplot(internal_assessment, aes(x = acceptance, y = impact, color = group)) +
  geom_point(position = pos, alpha = 0.5, size = internal_assessment$cost * 5) + 
  geom_label_repel(position = pos,aes(label = name)) +
  xlab("Acceptance") + ylab("Expected impact") +
  scale_y_continuous(breaks = c(1,2,3), labels = c("low","medium","high")) +
  labs(color = "Group") +
  scale_color_brewer(palette = "Dark2", type = "div") +
  theme_minimal() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        axis.title = element_text(size = 18),
        title = element_text(size = 18)) +
  guides(colour = guide_legend(override.aes = list(size=8)))
```

## List of measures

The following list of measures was originally proposed to Eawag employees for evaluation in the survey conducted by flyaware. It was elaborated based on gathering a crowd-sourced list of measures first within a core group and then within the extended flyaware group. To gather an initial set of measures, initiatives at other institutions, such as ETH and EPFL were taken into account. We tried to ensure a mix of regulatory, incentive-based and informational measures.

# {#official}
### *Official Statement (Eawag Directive)*: 
Eawag officially states that low-carbon transport (e.g. train) is preferred over high-carbon transport (e.g. flying) in an Eawag directive.

# {#intranet}
### *Intranet page*: 
Eawag sets up an intranet page on climate friendly mobility. This webpage will provide useful information on:
  -International train travel such as on booking platforms (e.g. loco2.com, trainline.com and interrail.eu) that facilitate the booking process of train tickets within Europe
  - Video conference training
  - The compensation system of CO2 emissions from flying at Eawag

# {#database}
### *Improved Data Base*: 
The Eawag reporting database, which collects data on conference contributions, supervision of students etc., will be extended to collect data on mobility behavior. For each international business trip, information on the following will be collected:
  -Type of transportation used (flight, train, bus, etc.)
  - Purpose of business trip (conference, project meeting, etc.)
  - Points of departure and arrival

# {#role_models}
### *Highlight role models*: 
Many Eawag employees already show a high awareness of the impact of their mobility behavior. The communications department will highlight Eawag employees who actively reduce their business flights, or avoid them entirely (e.g. via intranet, poster exhibition, etc.), in order to promote role models and to take away the fear that professional development may suffer by avoiding or reducing business flights.

# {#appraisal}
### *Integrate in appraisal interviews*: 
Reflections on climate friendly mobility become a mandatory part of yearly appraisal interviews. Supervisors and employees will reflect on whether a conference and the resultant business flight(s) are truly necessary to advancing given projects and/or professional development of the employee.

# {#showcase_video}
### *Showcasing video conferences*: 
Eawag promotes the videoconferencing technology by showcasing it during larger events, such as in Eawag seminars and conferences hosted at Eawag.

# {#video_training}
### *Trainings for videoconferencing*: 
Regular trainings on the recently installed video conference equipment will be available for all staff and will be mandatory for all group leaders.

# {#video_default}
### *Default videoconferencing for guests*: 
Eawag makes it official policy that video conferencing is the default choice for guests who would otherwise need to fly in for meetings and/or seminars.

# {#train_travel_policy}
### *Train travel as official policy*: 
Eawag will recommend and state within the developed directive (see measure official statement) that flights to destinations that can be reached within 10 hours by train, or that can be reached using an overnight train option, should be avoided. The journey will be considered as working time.

# {#train_travel_compensate}
### *Compensation of train travel*: 
Possible additional costs for train travel compared to flying (e.g. higher ticket price, additional overnight stays, etc.) will be paid for by Eawag e.g. by re-allocating funds received through compensation of flying-related CO2 emissions.

# {#train_firstclasse}
### *Finance first class trains*: 
Moreover, first-class train rides might be considered in certain cases (e.g. ticket price max. 30% higher than a standard rate ticket) and will be financially supported by Eawag (e.g. by re-allocating funds received through compensations of flying###related CO2 emissions (see measure 6)) .

# {#internal_competition}
### *Internal reduction targets per department*: 
After having established a database of mobility behavior over three years (see measure improved database) all departments are requested to set internal reduction targets, e.g. a reduction of 20-50% over the following five years compared to the collected three-year average.

# {#double_compensation}
### *Double CO2 compensation*: 
To further support climate friendly travel options, the CO2 compensation fee for all flights will be doubled.

# {#compensate_all_flights}
### *Compensate all flights*: 
All flights, independent of the funding source, are collected in the reporting database (see measure improved database). Hence, even flights paid for through external funding (e.g. invitations by other research institutes, universities etc.) - which haven’t previously been considered - will also be compensated.

# {#internal_comp_xmas}
### *Internal competition, highlight at Xmas*: 
Results of the mobility behavior collected in the reporting database (see measure improved database) will be presented at the Eawag Christmas party. Departments with the largest relative reduction of CO2 emissions related to their mobility behavior will be highlighted and awarded (e.g. by a department excursion sponsored by Eawag) every year.

# {#overseas_restriction}
### *Restrictions of Overseas Flights*: 
Overseas flights cause by far the largest amount of CO2 emissions among Eawag business flights. Therefore, Eawag will restrict the number of overseas flights (except for field work):
  - Group Leaders, Department Heads and the Directorate: 1 overseas flight within 2 years
  - PhD students and Postdocs: 1 overseas flight within PhD / Postdoc project period
  - Other staff: 1 overseas flight within 4 years (comparable to length of PhD project)

# {#internal_cap}
### *Internal cap*: 
Every department gets an annual CO2 emissions budget (e.g. based on the average CO2 emissions patterns from previous years, as collected in the reporting database (see measure 3)), which can be allocated and traded (e.g. in the shape of certificates) among departments and individuals.

# {#progressive_cap}
### *Progressive cap*: 
The cap could be progressive over time: increasing slowly for the first few years and then more quickly later, so that departments can plan ahead, prepare and adjust themselves to the updated regulations.

## List of flyaware members

- Andrea Rösch
- Jonas Heiberg
- Karin Kiefer
- Mario Angst
- Philipp Staudacher


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```