---
title: "Empfohlene Massnahmen zur Reduktion von Flugreisen an der Eawag"
author: "Flyaware interest group"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
# css: my_css.css
link-citations: yes
subtitle: Final report to the directorate
bibliography: skeleton.bib
---

```{r setup data, include=FALSE}
source("read_in_clean_data.R")
source("utility_functions.R")
```

<a href="index.html">English Version</a>

# A call for action

> "It is critical to immediately begin reducing net CO2 emissions and to eliminate them to zero worldwide between 2040 and 2050 at the latest."
>
> `r tufte::quote_footer('--- Scientists for future')`

This statement from the ‘Scientists for Future’^[https://www.scientists4future.org/statement-en-und-es/]  initiative was signed by 44 out of 87 (50.6%) Eawag group leaders and department heads. The director and deputy director of Eawag have signed this statement as well. Thus, the majority of the Eawag leadership is in agreement with this statement and is aware of the urgency of taking measures to reduce emissions.

One of the most effective ways for scientific institutions to reduce their climate impact is a drastic reduction of business flights. More than half of greenhouse gas emissions (GHGs) of Eawag's sister institution ETH Zurich are caused by business trips. Of those **93%** are due to business flights ^[https://ethz.ch/flugreisen]. Technical alternatives to substantially reduce GHG emissions of airplanes are not foreseeable for the near future. It is thus of immediate concern to reduce the number of flights.

Eawag is a research institution recognized worldwide and therefore has the potential to act as a role model for other research institutes by taking on its institutional responsibility for further reducing CO2 emissions. The time to be a rolemodel is now - academia and the world in general is finally waking up to the climate impacts of flying ^[for a helpful FAQ and original 2015 petition of academics worldwide see https://academicflyingblog.wordpress.com/]

This report is a call to the Eawag leadership to act. We propose nine concrete sets of measures to reduce Eawag’s CO2 emissions related to business flights. To aid in decision-making, We provide a list of highly accepted and highly impactful measures as priority targets for implementation.

```{r fig-margin, fig.margin = TRUE, fig.cap = "Employees of Eawag overwhelmingly support Eawag taking measures to aid in reducing business flights", fig.width=3.5, fig.height=3.5, cache=TRUE, message=FALSE, echo=FALSE}
plot_var_dist("should_eawag_support_flyless",y_limit = max_n,plot_xlabs = TRUE, plot_legend = FALSE, fill_var = "none", graph_title = "Eawag should support flying less")
```

We conducted a survey among all Eawag employees, in which roughly half of all employees (`r nrow(responses)`) participated. Employees were asked to evaluate the set of measures we present in this report. Given the high response rate, our results are highly likely to adequately reflect the views of Eawag employees on the topic of flying and the proposed measures.

The scientific evidence for man-made climate change and its dangers is robust. We therefore ask you as (environmental) scientists and members of the directorate to overcome the knowledge – action gap and support change here at Eawag.

## What measures should Eawag take?

We suggest the following list of measures for Eawag to take right away. ^[The complete set of measures as presented to Eawag employees in the survey can be <a href="static/complete_list_measures.pdf">downloaded here</a>]

<details>
  <summary> <font size="5"> Official Statement (Eawag Directive) </font> </summary>
  
  Eawag officially states that low-carbon transport (e.g. train) is preferred over high-carbon transport (e.g. flying) in an Eawag directive. 
  
</details> <p>

<details>
  <summary> <font size="5"> Intranet page </font> </summary>
  
  Eawag sets up an intranet page on climate friendly mobility. This webpage will provide useful information on:
  
  1. International train travel such as on booking platforms (e.g. loco2.com, trainline.com and interrail.eu) that facilitate the booking process of train tickets within Europe
  
  2. Video conference training 
  
  3. The compensation system of CO2 emissions from flying at Eawag
</details> <p>

<details>
  <summary> <font size="5"> Showcasing video conferences </font> </summary>
  
  Eawag promotes the videoconferencing technology by showcasing it during larger events, such as in Eawag seminars and conferences hosted at Eawag.
  
</details> <p>

<details>
  <summary> <font size="5"> Train travel as official policy </font> </summary>
  
  Eawag will recommend and state within the developed directive (see measure official statement) that flights to destinations that can be reached within 10 hours by train, or that can be reached using an overnight train option, should be avoided. The journey will be considered as working time.
  
</details> <p>

<details>
  <summary> <font size="5"> Compensation of train travel </font> </summary>
  
  Eawag officially states that low-carbon transport (e.g. train) is preferred over high-carbon transport (e.g. flying) in an Eawag directive. 
  
</details> <p>

<details>
  <summary> <font size="5"> Compensate all flights </font> </summary>
  
  All flights, independent of the funding source, are collected in the reporting database (see measure improved database). Hence, even flights paid for through external funding (e.g. invitations by other research institutes, universities etc.) - which haven’t previously been considered - will also be compensated.
  
</details> <p>

<details>
  <summary> <font size="5"> Double CO2 compensation </font> </summary>
  
  To further support climate friendly travel options, the CO2 compensation fee for all flights will be doubled.
  
</details> <p>

<details>
  <summary> <font size="5"> Internal reduction targets </font> </summary>
  
  After having established a database of mobility behavior over three years (see measure improved database) all departments are requested to set internal reduction targets, e.g. a reduction of 20-50% over the following five years compared to the collected three-year average.
  
</details> <p>

<details>
  <summary> <font size="5"> Improved database </font> </summary>
  
  The Eawag reporting database, which collects data on conference contributions, supervision of students etc., will be extended to collect data on mobility behavior. For each international business trip, information on the following will be collected: 
  
  1. Type of transportation used (flight, train, bus, etc.) 
  
  2. Purpose of business trip (conference, project meeting, etc.) 
  
  3. Points of departure and arrival  
    
</details> <p>

One of the most effective ways to reduce GHG emissions from flying is the reduction of long-distance flights ^[see eg. Jäckle, S. Eur Polit Sci (2019) 18: 630. https://doi.org/10.1057/s41304-019-00220-6]. Thus, we strongly recommend to prioritize a measure acting on long-distance flights.

<details>
  <summary> <font size="5"> Limit longdistance-flights </font> </summary>
  
  Overseas flights cause by far the largest amount of CO2 emissions among Eawag business flights. Therefore, Eawag will restrict the number of overseas flights (except for field work): 
  
  1. Group Leaders, Department Heads and the Directorate: 1 overseas flight within 2 years 
  
  2. PhD students and Postdocs: 1 overseas flight within PhD / Postdoc project period 
  
  3. Other staff: 1 overseas flight within 4 years (comparable to length of PhD project)
  
</details> <p>



## Selection of measures

To select the measures we propose for prioritized implementation we categorized them based on their acceptance among Eawag employees and their expected impact in terms of reducing GHG emissions. Acceptance was split into three groups, based on the distribution of acceptance^[However, it needs to be kept in mind that no single measure was rejected by a majority of survey respondents] among Eawag employees^[The distribution of variables can be explored interactively at: https://marioangst.shinyapps.io/flyaware_survey/ 
Survey data and analytical procedures can be accessed at https://github.com/marioangst/flyaware_survey]. An overview over all variables can be <a href="Viz_outputs/all_all_small_multiples.pdf">downloaded here.</a>
Expected impact and expected costs are based on an estimate within the flyaware group and as such only represent an initial prior. We welcome further discussion on these points.

We suggest to categorize measures into four groups:

- High acceptance: These measures are universally supported by Eawag employees. They are mostly relatively low-cost. We recommend their immediate adoption.
- Very disputed: The most controversial, yet generally most impactful measures. We recommend to forego these for the moment, except for the measure on limiting long-distance flights, due to its impact and because no other measure acts directly on long-distance flights.
- Disputed, medium impact: Somewhat disputed, medium-impact measures. We recommend the adoption of these measures.
- Disputed, low impact: Somewhat disputed measures with relatively little impact. We do recommend not to prioritize these measures, except concerning the adoption of a measure improving the database as it is the foundation for some other measures.

```{r, fig.width = 9, fig.height = 5, fig.cap = "Categorization of measures proposed based on acceptance, expected impact and expected cost. Each circle represents a measure, sized by expected costs.", warning=FALSE, message=FALSE, echo=FALSE}
library(ggrepel)

pos <- position_jitter(0.2,0.2,seed = 1)

ggplot(internal_assessment, aes(x = acceptance, y = impact, color = group)) +
  geom_point(position = pos, alpha = 0.5, size = internal_assessment$cost * 5) + 
  geom_label_repel(position = pos,aes(label = name)) +
  xlab("Acceptance") + ylab("Expected impact") +
  scale_y_continuous(breaks = c(1,2,3), labels = c("low","medium","high")) +
  labs(color = "Group") +
  scale_color_brewer(palette = "Dark2", type = "div") +
  theme_minimal() +
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=12),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        axis.title = element_text(size = 18),
        title = element_text(size = 18)) +
  guides(colour = guide_legend(override.aes = list(size=8)))
```

### Complete initial List of measures

The complete list of measures originally proposed to Eawag employees for evaluation in the survey conducted by flyaware can be <a href="static/complete_list_measures.pdf">downloaded here</a>. It was elaborated based on gathering a crowd-sourced list of measures first within a core group and then within the extended flyaware group. To gather an initial set of measures, initiatives at other institutions, such as ETH and EPFL were taken into account. We tried to ensure a mix of regulatory, incentive-based and informational measures.

## List of flyaware members

- Andrea Rösch
- Jonas Heiberg
- Karin Kiefer
- Mario Angst
- Philipp Staudacher


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```